Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-11-01 manuscript/Manuscript.Rmd
textchanges gerko

Potentially, this approach could fulfil the needs for openly accessible data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. The quality of the synthetic data is highly dependent on the imputation models, and using flawed imputation models might bias subsequent analyses [@reiter2004simultaneous; @grund2021using; @jiang2021balancing]. Conversely, if the models used to synthesize the data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. To fully exploit the benefits of synthetic data, complicating aspects of creating high-quality synthetic data sets should be kept at a minimum, while keeping the accuracy of the models as high as possible.

@@

However, when creating synthetic data with `mice`, close attention should be paid to three distinct aspects of the data: incorporating the additional uncertainty that is due synthesizing (part of) the data, potential disclosure risk that remain after synthesizing the data, and the utility of the synthetic data. First, the procedure of generating multiple synthetic sets may seem overly complicated. We would like to emphasize that analyzing a single synthesized set, while perhaps unbiased, would underestimate the variance properties that are so important in drawing statistical inferences from finite data sets. After all, we are often not interested in the sample at hand, but aim to make inferences about the underlying data generating mechanism as reflected in the population. Properly capturing the uncertainty of synthetic data sets, just like with incomplete data sets, is therefore paramount. To achieve this, we adopted a bootstrap scheme in our simulations, to represent sampling from a superpopulation. However, when a single sample should be the reference, in the sense that the sample reflects the population one wishes to make inferences about, adjusted pooling rules are required that are similar to the procedure outlined in @vink2014pooling. The corresponding pooling rules have not been derived yet and the incorporation in data analysis workflows would require proper attention from developers alike.

@@

Third, the utility of the synthetic data should receive considerable scrutiny. After all, synthetic data is generated to allow third parties without access to the data to make accurate inference. If the quality of the synthetic data is very poor, such accurate inferences are impossible. We showed that CART proved to be an adequate method for the data at hand. However, it may well be that for some problems, using CART for imputation is not the optimal strategy. We already demonstrated that imputing deterministic relationships should not be based on CART. However, due to the variety in possible imputation methods that have proved to perform well in varying circumstances, providing a single solution to all imputation problems seems impossible. If required, imputers could build up their imputation models in complexity. Nevertheless, similarly to past research [e.g., @burgette_reiter_cart_2010; @doove_buuren_recursive_2014; @raab_practical_2016], we showed that CART provides adequate results in general, and therefore serves as a good starting point.

After imputation, the quality of the synthetic data can be assessed using various utility measures. The simplest way to assess data quality is through visual inspection. However, visualizing more than two dimensions quickly becomes cumbersome, which makes it hard to assess the accuracy of multivariate relations. Trying to predict whether the data is observed or synthetic allows to get an intuition of the quality of the data. This procedure has been advanced by @snoke2018general, who proposed a measure of distributional similarity based on the mean squared error of propensity scores indicating class belonging for both observed and synthetic data. Additionally, if one knows the analyses that will be carried out by the synthetic data users, it is possible to compare the results on the observed and the synthetic data, allowing to assess the accuracy of the synthetisation method. Both of these methods have been implemented in the `R`-package `synthpop` [@synthpop], who pioneered in the field of synthetic data generation in `R`.

However, although `synthpop` is capable to synthesize regardless of missing data, it cannot solve eventual missingness problems. If the data for synthetisation contains missing values, the missingness will be considered as a property of the data, and the corresponding synthetic data sets will contain missing values as well. If the goal is to impute and synthesize a data set, a two-step approach must be used, in which the missing values are imputed using a different package, like `mice`, after which `synthpop` can be used to synthesize the imputed data. Then, given $m$ multiple imputations and $r$ synthetisations, at least $m \times r$ synthetic data sets are in order, after which the data can be analysed and the analyses can be pooled using rules developed by @reiter2004simultaneous. However, such an approach would be computationally inefficient and practically cumbersome, because the architectures of both packages differ.

The appealing nature of `mice` therefore lays in its ability to solve both the missingness problem and create synthetic versions of the data. Eventually, the flexibility with `mice` is that both unobserved and observed data values could be synthesized at once, without the need for a two-step approach. Using `mice`, $m$ synthetic sets would sufficient. However, as of today, no pooling rules for one-step imputation of missingness and synthetisation have been developed. The derivation of those would yield a massive theoretical improve to further reduce the burden of creating synthetic data sets. More practically, the `synthpop` package contains functionality that would be highly beneficial to users in `mice`, while `mice` is likely to contain functionality that would be beneficial to `synthpop`. Although efforts have been made to bridge gaps between the packages [e.g., @miceadds_3.11-6], further improvements could be made in that respect. Specifically, `synthpop` contains a wide variety of functions that can be used to assess the quality of synthetic data [e.g., @snoke2018general], while `mice` contains methods to impute multilevel data. Being able to use such functionality interchangeably would greatly benefit the public.

The ease of use when synthesizing data with `mice` in `R`, together with the validity of inferences obtained through this procedure opens up a wealth of possibilities for data dissemination and further research of initially private data.

>>>>

Potentially, this approach could fulfill the needs for openly accessible data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. The quality of the synthetic data is highly dependent on the imputation models, and using flawed imputation models might bias subsequent analyses [@reiter2004simultaneous; @grund2021using; @jiang2021balancing]. Conversely, if the models used to synthesize the data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. To fully exploit the benefits of synthetic data, complicating aspects of creating high-quality synthetic data sets should be kept at a minimum, while keeping the accuracy of the models as high as possible.

@@

However, when creating synthetic data with `mice`, close attention should be paid to three distinct aspects of the data: incorporating the additional uncertainty that is due to synthesizing (part of) the data, potential disclosure risks that remain after synthesizing the data, and the utility of the synthetic data. First, the procedure of generating multiple synthetic sets may seem overly complicated. We would like to emphasize that analyzing a single synthesized set, while perhaps unbiased, would underestimate the variance properties that are so important in drawing statistical inferences from finite data sets. After all, we are often not interested in the sample at hand, but aim to make inferences about the underlying data generating mechanism as reflected in the population. Properly capturing the uncertainty of synthetic data sets, just like with incomplete data sets, is therefore paramount. To achieve this, we adopted a bootstrap scheme in our simulations, to represent sampling from a superpopulation. However, when a single sample should be the reference, in the sense that the sample reflects the population one wishes to make inferences about, adjusted pooling rules are required that are similar to the procedure outlined in @vink2014pooling. The corresponding pooling rules have not been derived yet and the incorporation in data analysis workflows would require proper attention from developers alike.

@@

Third, the utility of the synthetic data should receive considerable scrutiny. After all, synthetic data is generated to allow third parties without access to the data to draw valid inferences. If the quality of the synthetic data is very poor, the validity of inferences is at stake. We showed that CART proved to be an adequate method for the data at hand. However, it may well be that for some problems, using CART for imputation is not the optimal strategy. We already demonstrated that imputing deterministic relationships should not be based on CART. However, due to the variety in possible imputation methods that have proved to perform well in varying circumstances, providing a single solution to all synthesis problems seems impossible. If required, imputers could build up their synthesis models in complexity. Nevertheless, similarly to past research [e.g., @burgette_reiter_cart_2010; @doove_buuren_recursive_2014; @raab_practical_2016], we showed that CART provides adequate results in general, and therefore serves as a good starting point.

After imputation, the quality of the synthetic data can be assessed using various utility measures. The simplest way to assess data quality is through visual inspection. However, visualizing more than two dimensions quickly becomes cumbersome, which makes it hard to assess the accuracy of multivariate relations. Trying to predict whether the data is observed or synthetic allows to get an intuition of the quality of the synthesis. This procedure has been advanced by @snoke2018general, who proposed a measure of distributional similarity based on the mean squared error of propensity scores indicating class belonging for both observed and synthetic data. Additionally, if one knows the analyses that will be carried out by the synthetic data users, it is possible to compare the results on the observed and the synthetic data, allowing to assess the accuracy of the synthetisation method. Both of these methods have been implemented in the `R`-package `synthpop` [@synthpop], who pioneered in the field of synthetic data generation in `R`.

While `synthpop` is capable to synthesize regardless of missing data, it is not developed to solve missingness problems. If the data for synthetisation contains missing values, the missingness will be considered as a property of the data, and the corresponding synthetic data sets will contain missing values as well. If the goal is to solve for the missingness and synthesize the data set, a two-step approach must be used, in which the missing values are imputed using a different package, like `mice`. After this, `synthpop` can be used to synthesize the imputed data. Then, given $m$ multiple imputations and $r$ synthetisations, at least $m \times r$ synthetic data sets are in order, after which the data can be analysed and the analyses can be pooled using rules developed by @reiter2004simultaneous. However, such an approach would be computationally inefficient and practically cumbersome, because the architectures of both packages differ.

The appealing nature of `mice` therefore lays in its ability to solve both the missingness problem and create synthetic versions of the data. Eventually, the flexibility with `mice` is that both unobserved and observed data values could be synthesized at once, without the need for a two-step approach. Using `mice`, $m$ synthetic sets would be sufficient. However, as of today, no pooling rules for one-step imputation of missingness and synthetisation have been developed. The derivation of those would yield a massive theoretical improvement to further reduce the burden of creating synthetic data sets.

That said, the fields of synthetic data and multiple imputation are very much in development. Although efforts have been made to bridge gaps between these fields, such as in the `miceadds` package [@miceadds_3.11-6] which links functionality between between `synthpop` and `mice`, further improvements could be made in that respect. Specifically, `synthpop` contains a comprehensive suite of functions that can be used to assess the quality of synthetic data [e.g., @snoke2018general], while `mice` contains methodology to allow for flexible synthesis of intricate data structures, such as multilevel data. Being able to use such functionality interchangeably would greatly benefit data analysist.

The added possibility of synthesizing data with `mice` in `R`, together with the validity of inferences obtained through this procedure opens up a wealth of possibilities for data dissemination and further research of initially private data to applied `R` users.

### 2021-11-01 manuscript/Manuscript.tex
textchanges gerko

Potentially, this approach could fulfil the needs for openly accessible
data, without running into barriers with regard to privacy and
confidentiality constraints. However, there is no such thing as a free
lunch: data collectors have to put effort in creating high-quality
synthetic data. The quality of the synthetic data is highly dependent on
the imputation models, and using flawed imputation models might bias
subsequent analyses
\citep{reiter2004simultaneous, grund2021using, jiang2021balancing}.
Conversely, if the models used to synthesize the data are able to
preserve the relationships between the variables as in the original
data, the synthetic data can be nearly as informative as the observed
data. To fully exploit the benefits of synthetic data, complicating
aspects of creating high-quality synthetic data sets should be kept at a
minimum, while keeping the accuracy of the models as high as possible.

@@

However, when creating synthetic data with \texttt{mice}, close
attention should be paid to three distinct aspects of the data:
incorporating the additional uncertainty that is due synthesizing (part
of) the data, potential disclosure risk that remain after synthesizing
the data, and the utility of the synthetic data. First, the procedure of
generating multiple synthetic sets may seem overly complicated. We would
like to emphasize that analyzing a single synthesized set, while perhaps
unbiased, would underestimate the variance properties that are so
important in drawing statistical inferences from finite data sets. After
all, we are often not interested in the sample at hand, but aim to make
inferences about the underlying data generating mechanism as reflected
in the population. Properly capturing the uncertainty of synthetic data
sets, just like with incomplete data sets, is therefore paramount. To
achieve this, we adopted a bootstrap scheme in our simulations, to
represent sampling from a superpopulation. However, when a single sample
should be the reference, in the sense that the sample reflects the
population one wishes to make inferences about, adjusted pooling rules
are required that are similar to the procedure outlined in
\citet{vink2014pooling}. The corresponding pooling rules have not been
derived yet and the incorporation in data analysis workflows would
require proper attention from developers alike.

@@

Third, the utility of the synthetic data should receive considerable
scrutiny. After all, synthetic data is generated to allow third parties
without access to the data to make accurate inference. If the quality of
the synthetic data is very poor, such accurate inferences are
impossible. We showed that CART proved to be an adequate method for the
data at hand. However, it may well be that for some problems, using CART
for imputation is not the optimal strategy. We already demonstrated that
imputing deterministic relationships should not be based on CART.
However, due to the variety in possible imputation methods that have
proved to perform well in varying circumstances, providing a single
solution to all imputation problems seems impossible. If required,
imputers could build up their imputation models in complexity.
Nevertheless, similarly to past research
\citep[e.g.,][]{burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016},
we showed that CART provides adequate results in general, and therefore
serves as a good starting point.

After imputation, the quality of the synthetic data can be assessed
using various utility measures. The simplest way to assess data quality
is through visual inspection. However, visualizing more than two
dimensions quickly becomes cumbersome, which makes it hard to assess the
accuracy of multivariate relations. Trying to predict whether the data
is observed or synthetic allows to get an intuition of the quality of
the data. This procedure has been advanced by \citet{snoke2018general},
who proposed a measure of distributional similarity based on the mean
squared error of propensity scores indicating class belonging for both
observed and synthetic data. Additionally, if one knows the analyses
that will be carried out by the synthetic data users, it is possible to
compare the results on the observed and the synthetic data, allowing to
assess the accuracy of the synthetisation method. Both of these methods
have been implemented in the \texttt{R}-package \texttt{synthpop}
\citep{synthpop}, who pioneered in the field of synthetic data
generation in \texttt{R}.

However, although \texttt{synthpop} is capable to synthesize regardless
of missing data, it cannot solve eventual missingness problems. If the
data for synthetisation contains missing values, the missingness will be
considered as a property of the data, and the corresponding synthetic
data sets will contain missing values as well. If the goal is to impute
and synthesize a data set, a two-step approach must be used, in which
the missing values are imputed using a different package, like
\texttt{mice}, after which \texttt{synthpop} can be used to synthesize
the imputed data. Then, given \(m\) multiple imputations and \(r\)
synthetisations, at least \(m \times r\) synthetic data sets are in
order, after which the data can be analysed and the analyses can be
pooled using rules developed by \citet{reiter2004simultaneous}. However,
such an approach would be computationally inefficient and practically
cumbersome, because the architectures of both packages differ.

The appealing nature of \texttt{mice} therefore lays in its ability to
solve both the missingness problem and create synthetic versions of the
data. Eventually, the flexibility with \texttt{mice} is that both
unobserved and observed data values could be synthesized at once,
without the need for a two-step approach. Using \texttt{mice}, \(m\)
synthetic sets would sufficient. However, as of today, no pooling rules
for one-step imputation of missingness and synthetisation have been
developed. The derivation of those would yield a massive theoretical
improve to further reduce the burden of creating synthetic data sets.
More practically, the \texttt{synthpop} package contains functionality
that would be highly beneficial to users in \texttt{mice}, while
\texttt{mice} is likely to contain functionality that would be
beneficial to \texttt{synthpop}. Although efforts have been made to
bridge gaps between the packages \citep[e.g.,][]{miceadds_3.11-6},
further improvements could be made in that respect. Specifically,
\texttt{synthpop} contains a wide variety of functions that can be used
to assess the quality of synthetic data
\citep[e.g.,][]{snoke2018general}, while \texttt{mice} contains methods
to impute multilevel data. Being able to use such functionality
interchangeably would greatly benefit the public.

The ease of use when synthesizing data with \texttt{mice} in \texttt{R},
together with the validity of inferences obtained through this procedure
opens up a wealth of possibilities for data dissemination and further
research of initially private data.

>>>>

Potentially, this approach could fulfill the needs for openly accessible
data, without running into barriers with regard to privacy and
confidentiality constraints. However, there is no such thing as a free
lunch: data collectors have to put effort in creating high-quality
synthetic data. The quality of the synthetic data is highly dependent on
the imputation models, and using flawed imputation models might bias
subsequent analyses
\citep{reiter2004simultaneous, grund2021using, jiang2021balancing}.
Conversely, if the models used to synthesize the data are able to
preserve the relationships between the variables as in the original
data, the synthetic data can be nearly as informative as the observed
data. To fully exploit the benefits of synthetic data, complicating
aspects of creating high-quality synthetic data sets should be kept at a
minimum, while keeping the accuracy of the models as high as possible.

@@

However, when creating synthetic data with \texttt{mice}, close
attention should be paid to three distinct aspects of the data:
incorporating the additional uncertainty that is due to synthesizing
(part of) the data, potential disclosure risks that remain after
synthesizing the data, and the utility of the synthetic data. First, the
procedure of generating multiple synthetic sets may seem overly
complicated. We would like to emphasize that analyzing a single
synthesized set, while perhaps unbiased, would underestimate the
variance properties that are so important in drawing statistical
inferences from finite data sets. After all, we are often not interested
in the sample at hand, but aim to make inferences about the underlying
data generating mechanism as reflected in the population. Properly
capturing the uncertainty of synthetic data sets, just like with
incomplete data sets, is therefore paramount. To achieve this, we
adopted a bootstrap scheme in our simulations, to represent sampling
from a superpopulation. However, when a single sample should be the
reference, in the sense that the sample reflects the population one
wishes to make inferences about, adjusted pooling rules are required
that are similar to the procedure outlined in \citet{vink2014pooling}.
The corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

@@

Third, the utility of the synthetic data should receive considerable
scrutiny. After all, synthetic data is generated to allow third parties
without access to the data to draw valid inferences. If the quality of
the synthetic data is very poor, the validity of inferences is at stake.
We showed that CART proved to be an adequate method for the data at
hand. However, it may well be that for some problems, using CART for
imputation is not the optimal strategy. We already demonstrated that
imputing deterministic relationships should not be based on CART.
However, due to the variety in possible imputation methods that have
proved to perform well in varying circumstances, providing a single
solution to all synthesis problems seems impossible. If required,
imputers could build up their synthesis models in complexity.
Nevertheless, similarly to past research
\citep[e.g.,][]{burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016},
we showed that CART provides adequate results in general, and therefore
serves as a good starting point.

After imputation, the quality of the synthetic data can be assessed
using various utility measures. The simplest way to assess data quality
is through visual inspection. However, visualizing more than two
dimensions quickly becomes cumbersome, which makes it hard to assess the
accuracy of multivariate relations. Trying to predict whether the data
is observed or synthetic allows to get an intuition of the quality of
the synthesis. This procedure has been advanced by
\citet{snoke2018general}, who proposed a measure of distributional
similarity based on the mean squared error of propensity scores
indicating class belonging for both observed and synthetic data.
Additionally, if one knows the analyses that will be carried out by the
synthetic data users, it is possible to compare the results on the
observed and the synthetic data, allowing to assess the accuracy of the
synthetisation method. Both of these methods have been implemented in
the \texttt{R}-package \texttt{synthpop} \citep{synthpop}, who pioneered
in the field of synthetic data generation in \texttt{R}.

While \texttt{synthpop} is capable to synthesize regardless of missing
data, it is not developed to solve missingness problems. If the data for
synthetisation contains missing values, the missingness will be
considered as a property of the data, and the corresponding synthetic
data sets will contain missing values as well. If the goal is to solve
for the missingness and synthesize the data set, a two-step approach
must be used, in which the missing values are imputed using a different
package, like \texttt{mice}. After this, \texttt{synthpop} can be used
to synthesize the imputed data. Then, given \(m\) multiple imputations
and \(r\) synthetisations, at least \(m \times r\) synthetic data sets
are in order, after which the data can be analysed and the analyses can
be pooled using rules developed by \citet{reiter2004simultaneous}.
However, such an approach would be computationally inefficient and
practically cumbersome, because the architectures of both packages
differ.

The appealing nature of \texttt{mice} therefore lays in its ability to
solve both the missingness problem and create synthetic versions of the
data. Eventually, the flexibility with \texttt{mice} is that both
unobserved and observed data values could be synthesized at once,
without the need for a two-step approach. Using \texttt{mice}, \(m\)
synthetic sets would be sufficient. However, as of today, no pooling
rules for one-step imputation of missingness and synthetisation have
been developed. The derivation of those would yield a massive
theoretical improvement to further reduce the burden of creating
synthetic data sets.

That said, the fields of synthetic data and multiple imputation are very
much in development. Although efforts have been made to bridge gaps
between these fields, such as in the \texttt{miceadds} package
\citep{miceadds_3.11-6} which links functionality between between
\texttt{synthpop} and \texttt{mice}, further improvements could be made
in that respect. Specifically, \texttt{synthpop} contains a
comprehensive suite of functions that can be used to assess the quality
of synthetic data \citep[e.g.,][]{snoke2018general}, while \texttt{mice}
contains methodology to allow for flexible synthesis of intricate data
structures, such as multilevel data. Being able to use such
functionality interchangeably would greatly benefit data analysist.

The added possibility of synthesizing data with \texttt{mice} in
\texttt{R}, together with the validity of inferences obtained through
this procedure opens up a wealth of possibilities for data dissemination
and further research of initially private data to applied \texttt{R}
users.
